---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# sparseNumeric

<!-- badges: start -->
<!-- The R CMD check badge will be inserted here by:
# usethis::use_github_actions_badge("R-CMD-check")
-->
[![R-CMD-check](https://github.com/santhoshipen/sparseNumeric/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/santhoshipen/sparseNumeric/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->


The sparseNumeric package implements an S4 class sparse_numeric that stores
only the non-zero entries of a numeric vector along with their positions, while
tracking the full length of the vector. This representation is useful for
working with sparse numeric data efficiently.

The package provides:

- the sparse_numeric S4 class and validity checks
- coercion to and from base R numeric
- arithmetic operations (+, -, *) via sparse_add, sparse_sub,sparse_mult, and sparse_crossprod
- methods for mean(), L2 norm(), and standardize()
- custom show() and plot() methods for sparse vectors

## Installation

You can install the development version of sparseNumeric like so:

From a local clone of the repository:

```{r}
# in the sparseNumeric project directory
# devtools::install()
```

Once the repository is on GitHub, you can install with:

```{r}
# install.packages("devtools")
# devtools::install_github("santhoshipen/sparseNumeric")
```

## Example

This is a basic example which shows you how to solve a common problem:

Load the package:
```{r example}
library(sparseNumeric)
## basic example code

# Create a sparse_numeric object

x_dense <- c(0, 0, 3, 0, 5)
x <- as(x_dense, "sparse_numeric")
x

# Mean 
mean(x)

# L2 Norm 
norm(x)

# Standardization 
standardize(x)

# Arithmetic with another sparse vector
y_dense <- c(1, 0, 3, 0, 2)
y <- as(y_dense, "sparse_numeric")

as(x + y, "numeric")
as(x - y, "numeric")
as(x * y, "numeric")

# Plot overlapping non-zero positions
plot(x, y)
```

## Why README.Rmd?

Using README.Rmd allows you to include runnable R code chunks, which
automatically generate output and figures in the knitted README.md.

To update the README, run:

```{r}
# devtools::build_readme()
```

You can also embed plots:
```{r}
# plot(pressure)
```
